---
title: "PP249 PSet 3"
author: "James Supeyo Kibiro"
format: html
editor: visual
---

```{r}
# Load libraries

library(tidyverse)
library(estimatr)
library(ggplot2)
```

## Section 1: Regression / Covariate Conceptual Review

Setup: You work at the United Nations (UN) and have a dataset from a randomized controlled\
trial (RCT) ascertaining whether giving people free condoms for the entirety of their adolescence\
reduces STI rates measured when they turn 18. Participants were individuals living in rural parts\
of Free State (a province in South Africa). The dataset includes 10 variables. Y is a binary\
outcome (has an STI at age 18), T is a treatment indicator (given free condoms by the UN\
throughout adolescence), and X1 – X8 are a series of covariates that ostensibly predict STI\
rates. X1 is a factor variable (the village the individual lives in) and X2 through X8 are\
continuous. The dataset is called “data” in R.

1.  What would you type in R to find the proportion of individuals assigned to\
    treatment? Why would that information be helpful? What’s an advantage of setting\
    this proportion to 50%?

```{r}

```
